PARTIAL TRACKING VIA LINEAR PROGRAMMING

### 2017-04-18 paper/paper.tex
Added section on computational complexity

As stated above, some of these constraints are redundant and can be removed.
Indeed, we have $\BS{0} \leq \BS{x} \leq \BS{1}$, therefore we will always have
$\BS{A}^{\text{I}}\BS{x} \geq 0$ and $\BS{A}^{\text{I}}\BS{x} \geq 0$.
Furthermore, all but the last row of (\ref{eq:cxnlcon}) can be seen as
constructed from linear combinations of rows of (\ref{eq:cxnbalcon}) and the last
row of (\ref{eq:cxnlcon}) so we only require $\BS{A}^{\text{C}}_{K-2}\BS{x} = L$.
Finally we always have $\BS{x} \leq \BS{1}$ because of the constraint that there
be a maximum of $1$ incoming and outgoing connection from each node.

The complete LP to find the $L$ best disjoint paths through a lattice described
by node connections $\rho$ is
\[
    \min_{\BS{x}} \cLP^{T} \BS{x} 
\]
subject to
\[
    \BS{G}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{I}} \\
        \BS{A}^{\text{O}} \\
        -\BS{I}
    \end{bmatrix} \BS{x} \leq
    \begin{bmatrix}
        \BS{1} \\
        \BS{1} \\
        \BS{0}
    \end{bmatrix}
\]
\begin{equation}
    \BS{A}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{B}} \\
        \BS{A}^{\text{C}}_{K-2}
    \end{bmatrix} \BS{x} = 
    \begin{bmatrix}
        \BS{0} \\
        L
    \end{bmatrix}
\end{equation}

% TODO: move to complexity

%\section{An optimal method \label{sec:lppathsearch}}
%
%There is a way to find a set of paths over multiple frames ($K > 2$) having the
%lowest total cost if
%we restrict the search to exactly $L$ paths. Instead of indexing parameters by
%their frame number $k$, we make $k$ part of the parameter set so that it can be
%used by the distance function $\mathcal{D}$. Assume that over $K$ frames there
%are $M$ total parameter sets. In this context we will consider them as nodes in
%a graph. We define the vector $\BS{c} \in \mathbb{R}^{M^2}$
%where the entry $\BS{c}_{i + Mj} = \mathcal{D} \left( \theta_{i}, \theta_{j}
%\right)$. If we have a set of connections $\Gamma_{i,j}$ we can calculate the
%total cost of these connections by defining the vector
%\[
%    \BS{x}_{i + Mj} = \begin{cases}
%        1 & \text{there is a connection between }i\text{ and }j\\
%        0 & \text{otherwise}
%    \end{cases}
%\]
%and then forming the inner product
%\[
%    c_{\text{total}}=\left\langle \BS{c},\BS{x} \right\rangle
%\]
%Note that a node cannot be connected to itself. The question is how to find
%$\BS{x}^{\ast}$ so that $c_{\text{total}}$ is minimized. How do we
%constrain $\BS{x}$ to give us a solution to the partial tracking
%problem? Let us consider an example.
%
%In Figure~\ref{plot:simple_graph} we have an example of a simple graph or
%lattice. Such a graph represents a plausible partial tracking situation:
%vertically aligned nodes are parameter sets estimated from the same analysis
%frame and we would like to connect these parameter sets between frames. The
%numbers are indices of nodes in the graph and the edges (possible connections) between
%them are indicated by lines. Imagine that we would like to
%find the two shortest paths. We will now examine the resulting paths from two
%algorithms using different criteria for shortness.
%
%In Figure~\ref{plot:simple_graph_greedy_paths} we find the paths using an
%algorithm similar to Algorithms~\ref{alg:mq_peak_match} but search instead over a
%tensor of distances $C \in \mathbb{R}^{3 \times 4 \times 2}$ whose entry
%$C_{i,j,k}$ represents the cost of travelling on the path connecting the $i$th
%node in layer 0, the $j$th node in layer 1 and the $k$th node in layer 2. This
%cost is the sum of the Euclidean distances giving the lengths of the
%connections. This is the greedy method of searching for the best paths whose
%optimality criterion is to find the set of best paths containing the absolute
%best path. We see in Figure~\ref{plot:simple_graph_greedy_paths} that the
%absolute shortest path, $1 \rightarrow 4 \rightarrow 8$, is discovered, followed
%by the second shortest path not using the nodes of the first path, $2
%\rightarrow 5 \rightarrow 7$.
%
%\begin{figure}[!t]
%    \centering
%    \centerline{\includegraphics[width=\figwidthscale\textwidth]{plots/small_graph_ex.eps}}
%    \CaptionWithTitle{%
%        \input{plots/small_graph_ex.txt}%
%    }{\label{plot:simple_graph}}
%\end{figure}
%
%\begin{figure}[!t]
%    \centering
%    \centerline{\includegraphics[width=\figwidthscale\textwidth]{plots/small_graph_ex_greedy_paths.eps}}
%    \CaptionWithTitle{%
%        \input{plots/small_graph_ex_greedy_paths.txt}%
%    }{\label{plot:simple_graph_greedy_paths}}
%\end{figure}
%
%\subsection{$L$ shortest paths via linear programming}
%
%To find a set of paths minimizing the total cost, we instead search for total
%solutions $\BS{x}$ that describe all paths in the graph. Assume for now
%that we can guarantee that the entries of $\BS{x}$ will be either 0 or
%1. In order to simplify the indexing in this section, we denote the number of
%nodes in each frame as $N_{k}$ ($M = \sum_{l=0}^{K-1} N_{k}$). To find a set of constraints for our search, we consider the structure of a
%valid solution $\BS{x}^{\ast}$. To maintain that paths not overlap, a
%valid solution's interior nodes (those not in the first or last frame) are only allowed to have one edge entering ---
%coming from a node in a previous frame --- and one edge leaving
%--- going to a node in a successive frame. To translate this into a
%constraint, consider the node $i$ and its possible $R_{i}$ successive connecting
%nodes $j_{0} \dotsc j_{R_{i}-1}$. Define the vector\footnote{The superscript s
%stands for ``successive''.}
%\[
%    a^{\text{s},i}_{i + Mj_{r}} = \begin{cases}
%        1 & \forall j_{r} \in \left[ j_{0} \dotsc j_{R_{i}-1} \right] \\
%        0 & \text{otherwise}
%    \end{cases}
%\]
%As all the entries of $\BS{x}$ are either 0 or 1, we have
%\[
%    0 \leq \left\langle \BS{a}^{\text{s},i}, \BS{x} \right\rangle \leq 1
%\]
%so we can make this a constraint to ensure that a node has at most one path
%leaving. Similarly, if we consider the node $j$ and its possible $R_{j}$ previous connecting
%nodes $i_{0} \dotsc i_{R_{j}-1}$, the vector\footnote{The superscript q
%stands for ``previous''.}
%\[
%    a^{\text{q},j}_{i_{r} + Mj} \begin{cases}
%        1 & \forall i_{r} \in \left[ i_{0} \dotsc i_{R_{j}-1} \right] \\
%        0 & \text{otherwise}
%    \end{cases}
%\]
%constrains that node $j$ have only one path entering through the constraint
%\[
%    0 \leq \left\langle \BS{a}^{\text{q},j} , \BS{x} \right\rangle \leq 1
%\]
%A node on a path will also have an edge entering and an edge leaving. To
%translate this into a constraint, we define a vector that counts the number of
%edges entering a node and subtracts then the number of edges leaving a node. The
%result should always be 0 for an equal number of edges entering and exiting a
%node. If $r$ is the index of the node considered, the vector is simply
%\footnote{The superscript b stands for ``balanced''.}
%\[
%    \BS{a}^{\text{b},r} = \BS{a}^{\text{q},r} -
%    \BS{a}^{\text{s},r}
%\]
%and the constraint
%\[
%    \left\langle \BS{a}^{\text{b},r}, \BS{x} \right\rangle = 0
%\]
%Finally we want to constrain that there be only $L$ paths. We do this by
%noticing that if this is true, there will be $L$ edges between frames $k$ and
%$k+1$. We constrain the number of paths going from edges
%$\Gamma_{k}$ in frame $k$ to $\Gamma_{k+1}$ by forming the vector
%\footnote{The superscript c stands for ``connections''.}
%\[
%    \BS{a}^{\text{c},k} = \sum_{j \in \Gamma_{k}}
%    \BS{a}^{\text{s},j}
%\]
%and asserting the constraint
%\[
%    \left\langle \BS{a}^{\text{c},k} , \BS{x} \right\rangle = L
%\]
%The length of $\BS{x}$ is $M^{2}$ so the total size of all the
%constraints is not insignificant, but most entries in the constraint vectors will
%be 0 and therefore the resulting constraint matrices very sparse, so sparse
%linear algebra routines can be used in computations. Furthermore, the
%$\BS{a}^{\text{b}}$ and $\BS{a}^{\text{c}}$ constraints are derived from
%$\BS{a}^{\text{q}}$ and $\BS{a}^{\text{s}}$, so only the latter need to be
%stored.
%
%\begin{figure}[!t]
%    \centering
%    \centerline{\includegraphics[width=\figwidthscale\textwidth]{plots/small_graph_ex_lp_paths.eps}}
%    \CaptionWithTitle{%
%        \input{plots/small_graph_ex_lp_paths.txt}%
%    }{\label{plot:simple_graph_lp_paths}}
%\end{figure}
%
%The complete \textit{linear program (LP)} solving the $L$ shortest paths problem is then
%\[
%    \min_{\BS{x}} \left\langle \BS{c}, \BS{x} \right\rangle
%\]
%subject to
%\[
%    \BS{0} \leq
%    \begin{bmatrix}
%        \BS{A}_{\text{s}} \\
%        \BS{A}_{\text{q}}
%    \end{bmatrix} \BS{x}
%    \leq \BS{1}
%\]
%\[
%    \begin{bmatrix}
%        \BS{A}_{\text{b}} \\
%        \BS{A}_{\text{c}}
%    \end{bmatrix}
%    \BS{x}
%    =
%    \begin{bmatrix}
%        \BS{0} \\
%        L\BS{1}
%    \end{bmatrix}
%\]
%\[
%    \BS{0} \leq \BS{x} \leq \BS{1}
%\]
%where $\BS{A}_{\text{s}}$ is the matrix with
%$\BS{a}^{\text{s},m}$ as its rows for $m \in [0 \dotsc M-1]$ and
%$\BS{A}_{\text{q}}$ is the matrix with $\BS{a}^{\text{q},m}$ as
%its rows, etc.
%
%The solution of the two best paths using the LP formulation
%is shown in Figure~\ref{plot:simple_graph_lp_paths} and a comparison of the
%total costs is shown in Table~\ref{tab:greedy_lp_cost_compare}
%
%\begin{table}[!b]
%    \caption{\label{tab:greedy_lp_cost_compare} Comparison of total costs in
%    Figure~\ref{plot:simple_graph_lp_paths}}
%    \begin{center}
%        \begin{tabular}{c c}
%            Greedy & LP \\
%            \hline
%            \input{plots/small_graph_ex_greedy_cost.txt} &
%            \input{plots/small_graph_ex_lp_cost.txt} \\
%        \end{tabular}
%    \end{center}
%\end{table}
%
%The LP formulation is inspired by a multiple object tracking algorithm for video
%\cite{jiang2007linear}. A proof that the solution $\BS{x}^{\ast}$ will
%have entries equal to either $0$ or $1$ can be found in
%\cite[q.~167]{parker1988discrete}. The theoretical computational complexity of
%the linear program is polynomial in the number of variables, see
%\cite{karmarkar1984new} for a proof and the demonstration of a fast algorithm
%for finding its solution. In practice, to extract paths from the solution, we do
%not test equality with $0$ or $1$ but rather test if the solution vector's
%values are greater than some threshold. 
%
\subsection{Complexity}

The LP formulation of the $L$-best paths problem gives results equivalent to the
solution to the $L$-best paths problem proposed in \cite{wolf1989finding}. The
complexity of our algorithm is different.  Assuming we
use the algorithm in \cite{karmarkar1984new} to solve the LP, our program has a
complexity of $O(M^{7}B^{2})$ where $M$ is the number of nodes (parameter sets)
and $B$ is the number of bits used to represent each number in the input. The
complexity of the algorithm by Wolf in \cite{wolf1989finding} is equivalent to
the Viterbi algorithm for finding the single best path through a trellis whose
$k$th frame has $\binom{N_{k}}{L}\binom{N_{k+1}}{L}L!$ connections where $N_{k}$
and $N_{k+1}$ are the number of nodes in two consecutive frames of the original
lattice. Therefore, assuming a constant number $N$ of nodes in each frame, its
complexity is $O((\binom{N}{L}^{2}L!)^{2}T)$. If there are few nodes in each
frame and a small number of paths are searched, Wolf's formulation is superior
as its complexity increases linearly with the number of frames in the lattice.
On the other hand, if each frame has a large number of nodes or many paths are
searched, the LP formulation is superior.  Informally we have found this to
agree with reality --- both algorithms were tried when producing the figures in
Section~\ref{sec:mq_lp_compare_chirp}.  Indeed the Wolf formulation took
prohibitively long to compute when many paths were desired, as did the LP when
many frames were considered.

In the special case that only 1 shortest path is searched the Viterbi algorithm
\cite{forney1973viterbi} can be used that only requires on the order of $N^{2}T$
calculations \cite{rabiner1989tutorial} where $N$ is the number of nodes in each
frame and $T$ is the number of frames (assuming the same number of nodes in each
frame).

>>>>

As stated above, some of these constraints are redundant and can be removed.
Indeed, we have $\BS{0} \leq \BS{x} \leq \BS{1}$, therefore we will always have
$\BS{A}^{\text{I}}\BS{x} \geq 0$ and $\BS{A}^{\text{O}}\BS{x} \geq 0$.
Furthermore, all but the last row of (\ref{eq:cxnlcon}) can be seen as
constructed from linear combinations of rows of (\ref{eq:cxnbalcon}) and the last
row of (\ref{eq:cxnlcon}) so we only require $\BS{A}^{\text{C}}_{K-2}\BS{x} = L$.
Finally we always have $\BS{x} \leq \BS{1}$ because of the constraint that there
be a maximum of $1$ incoming and outgoing connection from each node.

The complete LP to find the $L$ best disjoint paths through a lattice described
by node connections $\rho$ is then
\[
    \min_{\BS{x}} \cLP^{T} \BS{x} 
\]
subject to
\[
    \BS{G}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{I}} \\
        \BS{A}^{\text{O}} \\
        -\BS{I}
    \end{bmatrix} \BS{x} \leq
    \begin{bmatrix}
        \BS{1} \\
        \BS{1} \\
        \BS{0}
    \end{bmatrix}
\]
\begin{equation}
    \BS{A}\BS{x} =
    \begin{bmatrix}
        \BS{A}^{\text{B}} \\
        \BS{A}^{\text{C}}_{K-2}
    \end{bmatrix} \BS{x} = 
    \begin{bmatrix}
        \BS{0} \\
        L
    \end{bmatrix}
\end{equation}
where $\BS{I}$ is the identity matrix.

\section{Complexity}

% TODO: Memory complexity

Here we will compare the complexity of the LP formulation of the best $L$ paths
search to the greedy McAulay-Quatieri method as well a combinatorial algorithm
proposed in \cite{wolf1989finding}.

Assuming the same number of nodes $N$ in each frame of the lattice, the 
search for the $l$th best path in the generalized McAulay-Quatieri
algorithm ($0 \leq l < L$) requires a search over $(N-l)^{K}$ possible paths.

The LP formulation of the $L$-best paths problem gives results equivalent to the
solution to the $L$-best paths problem proposed in \cite{wolf1989finding}. The
complexity of the algorithm by Wolf in \cite{wolf1989finding} is equivalent to
the Viterbi algorithm for finding the single best path through a trellis whose
$k$th frame has $\binom{N_{k}}{L}\binom{N_{k+1}}{L}L!$ connections where $N_{k}$
and $N_{k+1}$ are the number of nodes in two consecutive frames of the original
lattice. Therefore, assuming a constant number $N$ of nodes in each frame, its
complexity is $O((\binom{N}{L}^{2}L!)^{2}K)$.

The complexity of the algorithm presented here is polynomial in the number of
variables (the size of $\BS{x}$).  Assuming we use the algorithm in
\cite{karmarkar1984new} to solve the LP, our program has a complexity of
$O(P^{3.5}B^{2})$ where $B$ is the number of bits used to represent each number
in the input.  However, this bound is conservative considering the reported
complexity of modern algorithms.

For instance, the complexity of a log-barrier interior-point method is dominated
by solving the system of equations
\begin{equation}
    \label{eq:kkt}
    \begin{bmatrix}
        -\BS{D}\BS{G}^{T}\BS{G} & \BS{A}^{T} \\
        \BS{A} & \BS{0}
    \end{bmatrix}
    \begin{bmatrix}
        \BS{u} \\
        \BS{v}
    \end{bmatrix}
    =
    \begin{bmatrix}
        t\cLP + \BS{A}^{T}\BS{d} \\
        \BS{0}
    \end{bmatrix}
\end{equation}
some $10$s of times \cite[p.~590]{boyd2004convex}. 
Each iteration then takes ${\frac{2}{3}((K-1)N^{2} + (K-2)N)^{3}}$ flops (floating-point
operations) to solve ($\ref{eq:kkt}$) using a standard $LU$-decomposition
\cite[p.~98]{golub1996matrix}. As
$\BS{D}$ is a diagonal matrix, if
the number of nodes in each frame is $N$ for all frames, then
$\BS{D}\BS{G}^{T}\BS{G}$ will be a block-diagonal matrix made up of $K-1$ blocks
$\BS{B}_k \in \mathbb{R}^{N^{2}\times N^{2}}$. The system can then be solved in
\begin{multline*}
    \frac{2}{3}(K-1)N^{6} + 2(K-2)(K-1)N^{5} + \\
    2(K-2)^{2}(K-1)N^{4} + \frac{2}{3}(K-2)^{3}N^{3}
\end{multline*}
flops \cite[p.~675]{boyd2004convex}; this complexity is without exploiting the
sparsity of $\BS{A}$ nor the structure of $\BS{B}_k = \BS{D}_{k}\BS{C}$ --- the product
of some diagonal matrix $\BS{D}_{k}$ with an unchanging symmetric matrix
$\BS{C}$.
